---
title: "AAtesting"
format: html
editor: visual
---

Necessary Imports

```{r}
library(tidyverse)
library(ggplot2)
library(readr)
```

Read in Data

```{r}
data <- read_csv("AAtestingdata.csv")

data2 <- read_csv("AAtestingdata2.csv")
```

Clean out Unwanted Columns

```{r}
data <- data |>
  select(-categorical_indices, 
         -`Encoding Method`, 
         -n_folds, 
         -classification_report) |>
  filter(!is.na(`Weighted F1 Score`))
```

Create Datasets of Interest

```{r}

baselinedata <- data |>
  filter(
    Method == "baseline",
    `Use Archetypes` == FALSE,
    `Imbalance Ratio` == 0.15
  )

baselinedata <- baselinedata |>
  select(-`Imbalance Ratio`, 
         -`Archetype Proportion`, 
         -`Reintroduced Minority`, 
         -`Use Archetypes`
         )

archetypedata <- data |> 
  filter(
    Method == "smote",
    `Use Archetypes` == TRUE,
  )
```

## Archetype Proportion Testing

```{r}

ggplot(archetypedata, aes(x = `Archetype Proportion`, y = `Weighted F1 Score`)) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = "lm", color = "blue") +
  labs(
    title = "Impact of Archetype Proportion on F1 Score by Dataset",
    x = "Archetype Proportion",
    y = "Weighted F1 Score"
  )

```

```{r}
model <- lm(`Weighted F1 Score` ~ `Archetype Proportion`, data = archetypedata)
summary(model)
```

Allow for splitting based upon whether the data is categorical or not, and whether or not the data set is multi-nomial classification.

```{r}

dataset_info <- data.frame(
  Dataset = c("artificial_tree", "audiology", "balance_scale", "breast_cancer", "car", "chess", "crx", "diabetes", 
              "ecoli_5", "flare1", "glass", "heart_disease", "heart_failure", "hepatitis", "hill_valley", "ionosphere", 
              "iris", "lymphography", "mnist_test", "optdigits", "parkinsons", "seeds", "segmentation", "sonar", 
              "tic-tac-toe", "titanic", "treeData", "waveform", "wine", "zoo"),
  Has_Categorical = c(FALSE, TRUE, TRUE, TRUE, TRUE, TRUE, TRUE, FALSE, FALSE, TRUE, FALSE, TRUE, TRUE, TRUE,
                      FALSE, TRUE, FALSE, TRUE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, TRUE, TRUE, FALSE, FALSE, FALSE, TRUE),
  Is_Multiclass = c(TRUE, TRUE, TRUE, FALSE, TRUE, FALSE, FALSE, FALSE, TRUE, TRUE, TRUE, FALSE, FALSE, FALSE, 
                    FALSE, FALSE, TRUE, TRUE, TRUE, TRUE, FALSE, TRUE, TRUE, FALSE, FALSE, FALSE, TRUE, TRUE, TRUE, TRUE)
)

archetypedatasplit <- inner_join(archetypedata, dataset_info, by = "Dataset")
```

```{r}
subset_1 <- archetypedatasplit |> filter(Is_Multiclass == TRUE, Has_Categorical == TRUE)
subset_2 <- archetypedatasplit |> filter(Is_Multiclass == TRUE, Has_Categorical == FALSE)
subset_3 <- archetypedatasplit |> filter(Is_Multiclass == FALSE, Has_Categorical == TRUE)
subset_4 <- archetypedatasplit |> filter(Is_Multiclass == FALSE, Has_Categorical == FALSE)

model_1 <- lm(`Weighted F1 Score` ~ `Archetype Proportion`, data = subset_1)
model_2 <- lm(`Weighted F1 Score` ~ `Archetype Proportion`, data = subset_2)
model_3 <- lm(`Weighted F1 Score` ~ `Archetype Proportion`, data = subset_3)
model_4 <- lm(`Weighted F1 Score` ~ `Archetype Proportion`, data = subset_4)

cat("\n--- Multiclass + Categorical ---\n")
summary(model_1)

cat("\n--- Multiclass + No Categorical ---\n")
summary(model_2)

cat("\n--- Binary + Categorical ---\n")
summary(model_3)

cat("\n--- Binary + No Categorical ---\n")
summary(model_4)

```

```{r}
archetypeinteractions <- lm(`Weighted F1 Score` ~ `Archetype Proportion` * Is_Multiclass * Has_Categorical, data = archetypedatasplit)
summary(archetypeinteractions)
```

Across our results we can conclude that Archetype Proportion doesn't have a significant impact on Weighted F1 Score given Archetypes are used. We used runner,py to generate results across all data sets and various Archetype Proportions and Reintroduced Minority.

## Reintroduced Minority Testing

```{r}

model <- lm(
  `Weighted F1 Score` ~ `Reintroduced Minority` * Is_Multiclass * Has_Categorical,
  data = archetypedatasplit
)

summary(model)

```

Reintroducing more minority examples helps performance, especially in categorical datasets, but that benefit drops sharply when the data is also multiclass — possibly due to oversampling creating confusion in multi-class setups with sparse categorical features.

```{r}
archetypedata2 <- data |> 
  filter(
    Method == "smote",
  )
archetypedata2 <- inner_join(archetypedata2, dataset_info, by = "Dataset")
```

```{r}
model <- lm(`Weighted F1 Score` ~ `Use Archetypes`, data = archetypedata2)
summary(model)
```

Adjusted R-squared is minimal, but p-value is significant that Use Archetypes seems to lower F1 score on average by about 0.028 points.

### Testing for Data set types.

```{r}
sub_mc_cat    <- filter(archetypedata2, Is_Multiclass == TRUE,  Has_Categorical == TRUE)
sub_mc_nocat  <- filter(archetypedata2, Is_Multiclass == TRUE,  Has_Categorical == FALSE)
sub_bin_cat   <- filter(archetypedata2, Is_Multiclass == FALSE, Has_Categorical == TRUE)
sub_bin_nocat <- filter(archetypedata2, Is_Multiclass == FALSE, Has_Categorical == FALSE)

model_mc_cat    <- lm(`Weighted F1 Score` ~ `Use Archetypes`, data = sub_mc_cat)
model_mc_nocat  <- lm(`Weighted F1 Score` ~ `Use Archetypes`, data = sub_mc_nocat)
model_bin_cat   <- lm(`Weighted F1 Score` ~ `Use Archetypes`, data = sub_bin_cat)
model_bin_nocat <- lm(`Weighted F1 Score` ~ `Use Archetypes`, data = sub_bin_nocat)

cat("\n--- Multiclass + Categorical ---\n")
summary(model_mc_cat)

cat("\n--- Multiclass + No Categorical ---\n")
summary(model_mc_nocat)

cat("\n--- Binary + Categorical ---\n")
summary(model_bin_cat)

cat("\n--- Binary + No Categorical ---\n")
summary(model_bin_nocat)
```

Even after filtering by data set type, Archetypes still don't seem to improve performance as opposed to non-archetypal smote, and although relatively small (-0.0615), significantly decreasing performance in binary classifications tasks where categorical variables are present.

## Testing for Veracity of Results with Reintroduced Minority

```{r}
archetypedata3 <- archetypedata2 |>
  filter(`Reintroduced Minority` == 0.7 | is.na(`Reintroduced Minority`))
```

```{r}
model <- lm(`Weighted F1 Score` ~ `Use Archetypes`, data = archetypedata3)
summary(model)
```

```{r}
sub_mc_cat    <- filter(archetypedata3, Is_Multiclass == TRUE,  Has_Categorical == TRUE)
sub_mc_nocat  <- filter(archetypedata3, Is_Multiclass == TRUE,  Has_Categorical == FALSE)
sub_bin_cat   <- filter(archetypedata3, Is_Multiclass == FALSE, Has_Categorical == TRUE)
sub_bin_nocat <- filter(archetypedata3, Is_Multiclass == FALSE, Has_Categorical == FALSE)

model_mc_cat    <- lm(`Weighted F1 Score` ~ `Use Archetypes`, data = sub_mc_cat)
model_mc_nocat  <- lm(`Weighted F1 Score` ~ `Use Archetypes`, data = sub_mc_nocat)
model_bin_cat   <- lm(`Weighted F1 Score` ~ `Use Archetypes`, data = sub_bin_cat)
model_bin_nocat <- lm(`Weighted F1 Score` ~ `Use Archetypes`, data = sub_bin_nocat)

cat("\n--- Multiclass + Categorical ---\n")
summary(model_mc_cat)

cat("\n--- Multiclass + No Categorical ---\n")
summary(model_mc_nocat)

cat("\n--- Binary + Categorical ---\n")
summary(model_bin_cat)

cat("\n--- Binary + No Categorical ---\n")
summary(model_bin_nocat)
```

```{r}
model <- lm(`Weighted F1 Score` ~ `Use Archetypes` * Dataset, data = archetypedata3)
summary(model)
```

Even after adjusting for the effect of Minority Reintroduction, the implementation of Archetypes doesn't seem to show significant beneficial effect in an data set type.

```{r}

```
